# Manuscript figures

## Library

Required packages:

```{r}
required <- c(
  "readxl",
  "magrittr",
  "lubridate",
  "janitor",
  "tidyverse",
  "stringi",
  "stringr",
  "gtsummary",
  "zoo",
  "patchwork",
  "cluster",
  "factoextra",
  "fitdistrplus",
  "mgcv",
  "scam",
  "sf"
)
```

Installing those that are not installed yet:

```{r}
to_install <- required[!required %in% installed.packages()[, "Package"]]
if (length(to_install))
  install.packages(to_install)
```

Loading the packages for interactive use:

```{r}
invisible(sapply(required, library, character.only = TRUE))
```

## Constant

```{r}

user <- Sys.getenv("USER")

if (user == "nguyenpnt") {
  path2data <- paste0(
    "/Users/nguyenpnt/Library/CloudStorage/OneDrive-OxfordUniversityClinicalResearchUnit/Data/HFMD/cleaned/"
  )
}

# Name of cleaned HFMD data file
cases_file <- "hfmd_hcdc.rds"

# Name of cleaned sero file
sero_file <- "sero_EV-A71_1222-1223.rds"

# Name of cleaned viro file
viro_file <- "virology_cleaned.rds"

# Name of outpatient data file
ch1_outpatient <- "CH1_outpatient_HCM_cleaned.rds"

path2ms_fig <- "/Users/nguyenpnt/OUCRU/HFMD/manuscript/fig/"
```

## Functions

Function to read data

```{r}
readRDS2 <- function(x) readRDS(paste0(path2data, x))
```

Function for table 1

```{r}
calculate_prop_test <- function(data, variable, by, ...) {
  # subset to non-missing
  data <- tidyr::drop_na(data, dplyr::all_of(c(variable, by)))

  prop.trend.test(
    x = table(data[[variable]], data[[by]])[2, ], # get the second row (the positive row)
    n = table(data[[by]])
  ) |>
    broom::tidy()
}
```

Overide select funtion

```{r}
select <- dplyr::select
```

Function to create the cohort lines

```{r}
cohort_lines <- function(cohort_ages = seq(0, 5, by = 1),viro = FALSE){
  start_date <- as.Date("2023-01-01")
  end_date   <- as.Date("2023-12-31")

  ## generate the cohort line
  tseq <- seq(start_date, end_date, by = "day")
  cohort_lines <- lapply(cohort_ages, function(a0){
    data.frame(
      date = tseq,
      age  = a0 + as.numeric(tseq - start_date) / 365,
      cohort = as.factor(a0)
    )
  }) |> dplyr::bind_rows()
  
  color_cohort <- ifelse(viro == TRUE,"black","#80FFFFFF")

  ## create mid-opacity white of cohort lines before June
  cohort_lines <- cohort_lines |>
    mutate(trend = case_when(
      date < as.Date("2023-06-01") ~ color_cohort,
      date >= as.Date("2023-06-01") ~ "#FFFFFFFF",
    ))
  cohort_lines$group <- consecutive_id(cohort_lines$trend)
  cohort_lines <- head(do.call(rbind, by(cohort_lines, cohort_lines$group, rbind, NA)), -1)
  cohort_lines[, c("trend", "group")] <- lapply(cohort_lines[, c("trend", "group")], na.locf)
  cohort_lines[] <- lapply(cohort_lines, na.locf, fromLast = T)

  cohort_lines$trend <- factor(cohort_lines$trend,
                               levels = c(color_cohort,"#FFFFFFFF"))
  cohort_lines
}
```

Constrained algorithm for serological data

```{r}
age_profile_constrained_cohort2 <- function(data,
                                            age_values = seq(0, 15, le = 512),
                                            ci = .95,
                                            n = 100) {
  predict2 <- function(...)
    predict(..., type = "response") |> as.vector()
  
  age_profile <- function(data,
                          age_values = seq(0, 15, le = 512),
                          ci = .95) {
    model <- gam(pos ~ s(age), binomial, data)
    link_inv <- family(model)$linkinv
    df <- nrow(data) - length(coef(model))
    p <- (1 - ci) / 2
    model |>
      predict(list(age = age_values), se.fit = TRUE) %>%
      c(list(age = age_values), .) |>
      as_tibble() |>
      mutate(
        lwr = link_inv(fit + qt(p, df) * se.fit),
        upr = link_inv(fit + qt(1 - p, df) * se.fit),
        fit = link_inv(fit)
      ) |>
      dplyr::select(-se.fit)
  }
  
  
  age_profile_unconstrained <- function(data,
                                        age_values = seq(0, 15, le = 512),
                                        ci = .95) {
    data |>
      group_by(collection) |>
      group_map( ~ age_profile(.x, age_values, ci))
  }
  
  shift_right <- function(n, x) {
    if (n < 1)
      return(x)
    c(rep(NA, n), head(x, -n))
  }
  
  dpy <- 365 # number of days per year
  
  mean_collection_times <<- data |>
    group_by(collection) |>
    summarise(mean_col_date = mean(col_date2)) |>
    with(setNames(mean_col_date, collection))
  
  cohorts <- cumsum(c(0, diff(mean_collection_times))) |>
    divide_by(dpy * mean(diff(age_values))) |>
    round() |>
    map(shift_right, age_values)
  
  age_time <- map2(mean_collection_times,
                   cohorts,
                   ~ tibble(collection_time = .x, cohort = .y))
  
  age_time_inv <- age_time |>
    map( ~ cbind(.x, age = age_values)) |>
    bind_rows() |>
    na.exclude()
  
  data |>
    # Step 1:
    group_by(collection) |>
    group_modify( ~ .x |>
                    age_profile(age_values, ci) |>
                    mutate(across(c(fit, lwr, upr), ~ map(
                      .x, ~ rbinom(n, 1, .x)
                    )))) |>
    group_split() |>
    map2(age_time, bind_cols) |>
    bind_rows() |>
    unnest(c(fit, lwr, upr)) |>
    pivot_longer(c(fit, lwr, upr), names_to = "line", values_to = "seropositvty") |>
    # Step 2a:
    filter(cohort < max(age) - diff(range(mean_collection_times)) / dpy) |>
    group_by(cohort, line) |>
    group_modify(
      ~ .x %>%
        scam(seropositvty ~ s(collection_time, bs = "mpi"), binomial, .) |>
        predict2(list(collection_time = mean_collection_times)) %>%
        tibble(collection_time = mean_collection_times, seroprevalence  = .)
    ) |>
    ungroup() |>
    group_by(collection_time, line) |>
    group_modify( ~ .x |>
                    mutate(across(
                      seroprevalence, ~ gam(.x ~ s(cohort), betar) |>
                        predict2()
                    ))) |>  ### modified
    ungroup() |>
    pivot_wider(names_from = line, values_from = seroprevalence) |>
    group_by(collection_time) |>
    group_split()
}
```

## Data

Serological data:

```{r}
df_sero <- readRDS2(sero_file)
```

Virological data:

```{r}
df_viro <- readRDS2(viro_file) %>%
  mutate(pos = case_when(
    sero_gr == "EV-A71" ~ TRUE,
    sero_gr != "EV-A71" ~ FALSE),
    pos_cv = case_when(
      sero_gr == "CV-A6" ~ TRUE,
      sero_gr != "CV-A6" ~ FALSE),
    peak = case_when(
      month(admission_date) >= 6 & month(admission_date) <= 8 ~ "6/2023 - 8/2023",
      month(admission_date) >= 9 ~ "9/2023 - 12/2023"
    ),
    adm_week = as.Date(floor_date(admission_date, "week")),
    adm_month = as.Date(floor_date(admission_date, "month")),
    time = as.numeric(admission_date)
  )
```

CH1 outpatient data:

```{r}
ch1_outpatient <- readRDS2(ch1_outpatient)
```

Case report data:

```{r}
df23 <- cases_file |>
  readRDS2() |> 
  dplyr::select(dob, admission_date, district, gender, inout, severity, medi_cen) |> 
  filter(year(admission_date) == 2023) |> 
  filter_out(is.na(dob)) |> 
  unique() |> 
  mutate(across(severity, ~ case_when(is.na(.x) ~ "Undefined",
                                      severity %in% c("3", "4") ~ "3 or 4",
                                      .default = severity) |> as.factor()),
         across(medi_cen, ~ tolower(stri_trans_general(.x, "Latin-ASCII"))),
         across(medi_cen,
                ~ case_when(.x == "benh vien nhi dong 1" ~ "Children's hospital 1",
                            .x == "benh vien benh nhiet doi tphcm" ~
                              "Hospital of Tropical Diseases",
                            .x == "benh vien nhi dong 2" ~ "Children's hospital 2",
                            .x == "benh vien nhi dong thanh pho" ~
                              "City Children's hospital", .default = .x)),
         age1     = interval(dob, admission_date) / years(1),
         time     = as.numeric(admission_date),
         adm_week = as.Date(floor_date(admission_date, "week")),
         age      = round(age1))
```

HCMC map

```{r}
vn_qh <- st_read(dsn = paste0(path2data, "gadm41_VNM.gpkg"), layer = "ADM_ADM_2")

vn_qh1 <- vn_qh %>%
  clean_names() %>%     ## cho thành chữ thường
  filter(
    str_detect(
      name_1,
      "Hồ Chí Minh"
    )
  )
qhtp <- vn_qh1[-c(14,21),]

qhtp$geom[qhtp$varname_2 == "Thu Duc"] <- vn_qh1[c("21","24","14"),] %>%
  st_union()

qhtp <- qhtp %>% st_cast("MULTIPOLYGON")

qhtp$varname_2 <- stri_trans_general(qhtp$varname_2, "latin-ascii") %>%
  tolower() %>%
  str_remove("district") %>%
  trimws(which = "both")

qhtp$nl_name_2 <- c("BC","BTân","BT","CG","CC","GV",
                    "HM","NB","PN","1","10","11","12",
                    "3","4","5","6","7","8","TB",
                    "TP","TĐ")

## children hospital 1 coord

tdnd1 <- data.frame(long = 106.6702,
                    lat  = 10.7686) %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326)
```

## Results

### K-mean

Using k-mean clustering algorithm to separate 2 waves

```{r}
set.seed(123)

kmeans_result_1 <- kmeans(df23[, c("age1", "time")], centers = 2, nstart = 25)

df23$wave <- factor(kmeans_result_1$cluster,
                    labels = c("1st", "2nd"),
                    levels = c(2, 1))

## show the time point that seperate two peaks

time_cut_1st <- df23 |>
  filter(wave == "1st") |>
  pull(time) |>
  max()

time_cut_2nd <- df23 |>
  filter(wave == "2nd") |>
  pull(time) |>
  min()

time_cut <- mean(time_cut_1st, time_cut_2nd) |> as.numeric()
time_cut
```

So I selected time points at `r time_cut` to seperate two peaks

```{r}
sup1 <- df23 |>
  filter(age1 <= 15) |>
  ggplot(aes(x = admission_date, y = age1, color = wave)) +
  geom_point(size = .5) +
  scale_x_datetime(date_breaks = "1 month" , date_labels = "%b") +
  theme_minimal() +
  geom_vline(xintercept = as.Date(time_cut), linewidth = .2) +
  labs(x = "Admission date (2023)", y = "Age (years)") +
  guides(color = guide_legend(override.aes = list(size = 8), title = "Wave")) +
  theme(
    axis.title.x = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    axis.title.y = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 18),
    legend.title = element_text(size = 18)
  )
```

```{r}
ggsave(
  paste0(path2ms_fig,"sup1.svg"),
  plot = sup1,
  width = 15,
  height = 9,
  dpi = 500
)
```

### Table 1

```{r}
tab_age <-
  df23 |>
  tbl_summary(
    by = wave,
    label = list(age1 ~ "Age (Years)",
                 gender ~ "Gender"),
    digits = list(all_continuous() ~ c(2, 2), all_categorical() ~ c(0, 2)),
    include = c(age1,gender)
  ) |>
  remove_row_type(type = "missing")

tab_sev <-
  df23 |>
  filter(severity != "Undefined") |>
  dplyr::select(wave, severity)  |>
  mutate(rn = row_number()) |>
  tidyr::spread(severity, severity) |>
  mutate(across(-c(wave, rn), ~ ifelse(is.na(.), 0, 1))) |>
  dplyr::select(-rn) |>
  tbl_summary(
    by = wave,
    statistic = list(all_categorical() ~ "{n} / {N} ({p}%)"),
    digits = all_categorical() ~ c(0, 0, 2)
  ) |>
  modify_table_body(~ .x |>
                      mutate(row_type = "level", variable = "wave") %>%
                      {
                        bind_rows(tibble(
                          variable = "wave",
                          row_type = "label",
                          label = "Severity"
                        ),
                        .)
                      })

tab_inout <-
  df23 |>
  filter(inout %in% c("Inpatient", "Outpatient")) |>
  select(wave, inout) |>
  mutate(rn = row_number()) |>
  tidyr::spread(inout, inout) |>
  mutate(across(-c(wave, rn), ~ ifelse(is.na(.), 0, 1))) |>
  select(-rn) |>
  tbl_summary(
    by = wave,
    statistic = list(all_categorical() ~ "{n} / {N} ({p}%)"),
    digits = all_categorical() ~ c(0, 0, 2)
  ) |>
  modify_table_body(~ .x |>
                      mutate(row_type = "level", variable = "wave") %>%
                      {
                        bind_rows(tibble(
                          variable = "wave",
                          row_type = "label",
                          label = "Treatment"
                        ),
                        .)
                      })

final_table <- tbl_stack(list(tab_age, tab_sev, tab_inout), quiet = TRUE)

final_table
```

### Figure 1

::: {.panel-tabset .nav-pills}

#### Fig 1A

```{r}
f1a <- df23 |>
  group_by(adm_week,wave) |> 
  count() |> 
  ungroup() |> 
  ggplot(aes(x = adm_week, y = n,fill = wave))+
  geom_col(alpha = 0.6,color="black")+
  scale_x_date(date_breaks = "1 month",
               date_labels = "%b",
               name = "Admission date (2023)",
               limits = c(as.Date("2023-01-01"),as.Date("2023-12-31")))+
  # geom_vline(xintercept = as.Date(time_cut))+
  scale_fill_manual(values = c("#582C83FF","#FFC72CFF"))+
  theme_minimal()+
  labs(y = "Cases",tag = "A")+
  theme(axis.title.y = element_text(size = 18),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "bottom",
        plot.tag = element_text(face = "bold", size = 18),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 18),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 18))+
  guides(fill = guide_colourbar(barwidth=20),
         color = "none")
```

#### Fig 1B

```{r}
## aggregate number of admission cases overtime
df_smooth <- df23 |>
  mutate(age3 = floor(age1)) |>
  group_by(admission_date, age3) |>
  summarize(n = n(), .groups = "drop") |>
  complete(age3, admission_date, fill = list(n = 0)) |>
  mutate(
    age = as.numeric(age3),
    time = as.numeric(admission_date)   # convert to numeric scale
  )


## gam formular with tensor-product
fit_gau_bs <- gam(
  n ~ s(age, bs = "bs") +
    s(time, bs = "bs") +
    ti(age, time, bs = c("bs", "bs")) ,
  data = df_smooth,
  family = gaussian(link = "log"),
  method = "REML"
)

## create data for predict
age_val <- seq(0, 6, le = 365)

collection_date_val <- seq(min(df23$adm_week), max(df23$adm_week), le = 365)

new_data <- expand.grid(age = age_val, time = as.numeric(collection_date_val))
## predict
new_data$pred_gau_bs <- predict(fit_gau_bs, newdata = new_data, type = "response")

f1b <- ggplot(new_data) +
  geom_raster(aes(x = as.Date(time), y = age, fill = pred_gau_bs), interpolate = TRUE) +
  geom_line(
    data = cohort_lines(),
    aes(
      x = date,
      y = age,
      group = cohort,
      color = trend
    ),
    linewidth = 0.3,
    alpha = 0.4
  ) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  scale_fill_viridis_c(option = "turbo", n.breaks = 10) +
  theme_minimal() +
  labs(x = "Admission date",
       y = "Age (years)",
       fill = "Number of\nadmissions",
       tag = "B") +
  theme_minimal() +
  scale_y_reverse(name = "Age (years)",
                  lim = rev(c(-0.5, 6.5)),
                  breaks = seq(-1, 7)) +
  theme(
    axis.title.y = element_text(size = 18),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    plot.tag = element_text(face = "bold", size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = 18)
  ) +
  guides(fill = guide_colourbar(barwidth = 25), color = "none")
```

#### Fig 1C

```{r}
fit1 <- fitdist(df23 %>% filter(age1 > 0 &
                                  wave == "1st") %>% pull(age1), "lnorm")
fit2 <- fitdist(df23 %>% filter(age1 > 0 &
                                  wave == "2nd") %>% pull(age1), "lnorm")

x_vals <- seq(0, 7, le = 512)


fit_1p <- data.frame(
  wave = "1st",
  age = x_vals,
  density = dlnorm(
    x_vals,
    meanlog = fit1$estimate[1],
    sdlog = fit1$estimate[2]
  )
)

fit_2p <- data.frame(
  wave = "2nd",
  age = x_vals,
  density = dlnorm(
    x_vals,
    meanlog = fit2$estimate[1],
    sdlog = fit2$estimate[2]
  )
)

f1c <- ggplot(data = df23) +
  geom_histogram(aes(x = age1, fill = wave),
                 color = "white",
                 alpha = .5) +
  scale_fill_manual(values = c("#582C83FF", "#FFC72CFF")) +
  geom_line(data = rbind(fit_1p, fit_2p), aes(x = age, y = density * 5000)) +
  facet_wrap( ~ wave) +
  scale_x_reverse(
    limit = c(6, 0),
    breaks = seq(0, 6, by = 1),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    name = "Cases count",
    sec.axis = sec_axis(trans =  ~ . / 5000, name = "Density")
  ) +
  coord_flip() +
  theme_bw() +
  labs(x = "Age", y = "Density", tag = "C") +
  theme(
    axis.title.y = element_blank(),
    axis.ticks.x = element_blank(),
    # legend.position = "top",
    plot.tag = element_text(face = "bold", size = 18),
    axis.title.x = element_text(size = 18),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 18),
    legend.text = element_text(size = 18),
    strip.text = element_text(size = 18),
    legend.title = element_text(size = 18)
  )
```

:::

```{r}
#| fig-width: 15
#| fig-height: 9
#| out-width: "100%"

fig1 <- ((f1a | plot_spacer()) /
           (f1b | wrap_elements(full = f1c  +
                                  theme(
                                    plot.margin = margin(-40, 0, 67, 0)
                                  )))) +
  plot_layout(widths = c(1, 1))

ggsave(
  paste0(path2ms_fig,"fig1.svg"),
  plot = fig1,
  width = 15,
  height = 9,
  dpi = 500
)
fig1
```


### Fig 2

Case report CH1

```{r}
df_cases_ch1_23 <- df23 |>
  filter(medi_cen == "Children's hospital 1")
```

Outpatient admitted to CH1

```{r}
expected_age_cm <- ch1_outpatient |> 
  group_by(period,age_gr) |> 
  summarise(total_adm = sum(n),.groups = "drop") |>  
  group_by(period) |> 
  group_modify(~ gam(total_adm~s(age_gr),method = "REML",data = .) %>% 
                 predict(list(age_gr = seq(0,15,le = 512)),type="response") %>%
                 as.tibble() %>% 
                 mutate(age = seq(0,15,le = 512)))
```

Serological data CH1

```{r}
df_sero <- df_sero |>
  as_tibble() |>
  mutate(
    collection = id |>
      str_remove(".*-") |>
      as.numeric() |>
      divide_by(1e4) |>
      round(),
    col_date2 = as.numeric(col_date),
    across(pos, ~ .x > 0)
  )
```

Constraint algorithm

```{r}
# outttt <- age_profile_constrained_cohort2(df_sero)
# save(outttt,mean_collection_times,file = paste0(path2ms_fig,"sero.RData"))
load(paste0(path2ms_fig,"sero.RData"))
```

::: {.panel-tabset .nav-pills}

#### Fig 2A

```{r}
ouut <- list()
dat <- list()

for (i in 1:3) {
  dat[[i]] <- df_cases_ch1_23 |>
    filter(
      admission_date >= as.Date(mean_collection_times[i]) &
        admission_date <= as.Date(mean_collection_times[i + 1])
    )
}


expected_incidence <- map2(
  head(outttt, -1),
  tail(outttt, -1),
  ~ left_join(na.exclude(.x), na.exclude(.y), "cohort") |>
    mutate(
      attack = (fit.y - fit.x) / (1 - fit.x),
      date = as.Date(collection_time.y)
    )
) |>
  bind_rows() |>
  mutate(
    period = case_when(
      date <= as.Date("2023-04-30") ~ "12/2022 - 4/2023",
      date > as.Date("2023-04-30") &
        date <= as.Date("2023-08-31") ~ "4/2023 - 8/2023",
      date > as.Date("2023-08-31") ~ "8/2023 - 12/2023"
    )
  ) |>
  group_by(period) |>
  group_split() |>
  map( ~ left_join(.x, expected_age_cm, by = join_by(cohort == age, period))) %>%
  bind_rows(.id = "id") |>
  mutate(sp_gap = fit.y - fit.x, exp_inc = sp_gap * value)
```

First I estimate the standard error of difference in seroprevalence between 2 successive collection time points by the following formula:

$$
SE_{diff} = \sqrt{SE_1^2 + SE_2^2 - 2\rho . SE_1 . SE_2}
$$

$\rho$ is the assumed correlation of seroprevalence between two successive collection times points. Because we applied the constraint algorithm, I think we assumed two collection time points are dependence. So I choose $\rho = 1$

```{r}
rho = 1

expected_incidence <- expected_incidence |>
  mutate(
    se1 = (upr.x - lwr.x) / (2 * 1.96),
    se2 = (upr.y - lwr.y) / (2 * 1.96),
    se.diff = sqrt(se1^2 + se2^2 - 2 * rho * se1 * se2),
    lo_diff = sp_gap - 1.96 * se.diff,
    hi_diff = sp_gap + 1.96 * se.diff,
    hi_exp_inc = hi_diff * value,
    lo_exp_inc = lo_diff * value
  ) 

```

```{r}
fig2a <- expected_incidence |>
  ggplot() +
  geom_line(aes(x = cohort, y = exp_inc),linewidth = 1) +
  geom_ribbon(
    aes(x = cohort, y = exp_inc, ymin = lo_exp_inc, ymax = hi_exp_inc),
    alpha = 0.2,
    fill = "blue"
  ) +
  coord_cartesian(ylim = c(0, 2500)) +
  geom_col(
    data = dat %>%
      bind_rows(.id = "id") %>%
      filter(age1 > 0) |>
      group_by(id, age) |>
      count(),
    aes(x = age, y = n),
    color = "white",
    fill = "black",
    alpha = 0.2
  ) +
  facet_wrap( ~ factor(
    id,
    labels = c("12/2022 - 4/2023", "4/2023 - 8/2023", "8/2023 - 12/2023")
  )) +
  labs(y = "Number of cases", x = "Age (years)", tag = "A") +
  scale_x_continuous(position = "top")+
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 18),
    plot.tag = element_text(face = "bold", size = 18),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    legend.title = element_text(size = 18),
    strip.text = element_text(size = 18),
    panel.grid.minor.x = element_blank()
  )
```

#### Fig 2B

```{r}
fig2b <- outttt |>
  bind_rows() |>
  mutate(
    col_date = as.Date(collection_time),
    col_lab = format(col_date, format = "%b %Y"),
    col_lab = factor(col_lab, levels = c(
      "Dec 2022", "Apr 2023", "Aug 2023", "Dec 2023"
    ))
  ) |>
  ggplot(aes(x = cohort, y = fit)) +
  geom_line() +
  geom_ribbon(aes(ymin = lwr, ymax = upr),
              fill = "blue",
              alpha = .3) +
  geom_point(data = df_sero |>
               mutate(col_lab = factor(
                 col_time, levels = c("Dec 2022", "Apr 2023", "Aug 2023", "Dec 2023")
               )),
             aes(x = age, y = pos),
             shape = "|") +
  facet_wrap(~ col_lab, nrow = 1) +
  labs(y = "Seroprevalence (%)", x = "Age (years)", tag = "B") +
  scale_y_continuous(labels = scales::label_percent(scale = 100),
                     limits = c(0, 1)) +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    legend.position = "none",
    plot.tag = element_text(face = "bold", size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    strip.text.x = element_text(size = 18)
  )
```

#### Fig 2C

```{r}
fig2c <- df_cases_ch1_23 %>%
  group_by(adm_week, wave) %>%
  count() %>%
  ggplot(aes(x = adm_week, y = n, fill = wave)) +
  geom_col(alpha = 0.6, color = "black") +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %Y",
    name = "Admission week",
    limits = c(as.Date("2023-01-01"), as.Date("2023-12-31"))
  ) +
  scale_y_continuous(limits = c(0, 900), breaks = seq(0, 900, le = 4)) +
  scale_fill_manual(values = c("#582C83FF", "#FFC72CFF")) +
  theme_minimal() +
  annotate(
    "rect",
    fill = "grey",
    xmin = as.Date(c(
      "2023-01-01", "2023-04-05", "2023-08-02", "2023-12-06"
    )),
    xmax = as.Date(c(
      "2023-01-15", "2023-04-26", "2023-08-30", "2023-12-27"
    )),
    ymin = 0,
    ymax = Inf,
    alpha = .5
  ) +
  labs(y = "Cases",tag = "C") +
  theme(
    axis.title.y = element_text(size = 18),
    axis.title.x = element_text(size = 18),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    plot.tag = element_text(face = "bold", size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = 18)
  ) +
  guides(fill = guide_colourbar(barwidth = 20), color = "none")
```

:::

```{r}
#| fig-width: 20
#| fig-height: 16
#| out-width: "100%"

fig2 <- (fig2a /
               fig2b /
               fig2c) +
  plot_layout(heights = c(1, 1, 1))

ggsave(
  paste0(path2ms_fig,"fig2.svg"),
  plot = fig2,
  width = 20,
  height = 16,
  dpi = 500
)

fig2
```


heatmap for ch1 (sup fig 3)

```{r}
df_cases_count_ch1_23 <- df_cases_ch1_23 |>
  group_by(age, time) |>
  count() |>
  ungroup()

fit_gau_bs_ch1 <- gam(
  n ~ s(age, bs = "bs") +
    s(time, bs = "bs") +
    ti(age, time, bs = c("bs", "bs")) ,
  data = df_cases_count_ch1_23,
  family = gaussian(link = "log"),
  method = "REML"
)

age_val <- seq(0, 6, le = 365)

collection_date_val <- seq(min(df23$adm_week), max(df23$adm_week), le = 365)

new_data2 <- expand.grid(age = age_val, time = as.numeric(collection_date_val))

new_data2$pred_gau_bs_ch1 <- predict(fit_gau_bs_ch1, newdata = new_data2, type = "response")
```

```{r}
sup3b <- ggplot(new_data2) +
  geom_raster(aes(x = as.Date(time), y = age, fill = pred_gau_bs_ch1), interpolate = TRUE) +
  geom_line(
    data = cohort_lines(),
    aes(
      x = date,
      y = age,
      group = cohort,
      color = trend
    ),
    linewidth = 0.3,
    alpha = 0.4
  ) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  scale_fill_viridis_c(option = "turbo", n.breaks = 10) +
  theme_minimal() +
  labs(x = "Admission Week",
       y = "Age (years)",
       fill = "Number of\nadmissions",
       tag = "B") +
  theme_minimal() +
  scale_y_reverse(name = "Age (years)",
                  lim = rev(c(-0.5, 6.5)),
                  breaks = seq(-1, 7)) +
  theme(
    axis.title.y = element_text(size = 18),
    axis.title.x = element_text(size = 18),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    plot.tag = element_text(face = "bold", size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = 18)
  ) +
  guides(fill = guide_colourbar(barwidth = 25), color = "none")
```

### Fig 3

```{r}
per_serotype <- df_viro %>%
  group_by(adm_month) %>%
  count(sero_gr) %>%
  mutate(total = sum(n),
         per = n/total) %>%
  ggplot(aes(x = adm_month,y=per,fill = sero_gr))+
  geom_bar(position="fill", stat="identity")+
  scale_x_date(date_breaks = "1 month", date_labels = "%b",
               limits = c(as.Date("2023-01-01"),as.Date("2023-12-31")),
               name = "Collection month")+
  scale_y_continuous(labels = scales::label_percent())+
  labs(fill = "Serotype group",y = "Percentage (%)",tag = "B")+
  theme_minimal()+
  theme(axis.title.y = element_text(size = 18),
        axis.title.x = element_text(size = 18),
        axis.ticks.x = element_blank(),
        legend.position = "bottom",
        plot.tag = element_text(face = "bold", size = 18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 18))
```

Model perecentage of EV-A71 as a function of age and time

```{r}
## model ev-a71 percentage as a function of time and age

model_viro_age_time <- gam(pos ~ s(age_adm) + s(time) + ti(age_adm,time),
                           family = binomial,method = "REML",data = df_viro)

collection_date_val <- seq(min(df_viro$time),
                           max(df_viro$time), le = 365)

new_data2 <- expand.grid(age_adm = age_val,
                         time = as.numeric(collection_date_val))

new_data2$pred <- predict(model_viro_age_time, newdata = new_data2, type = "response")
```

Reconstruct epicurve

Model perecentage of EV-A71 as a function of age and time

```{r}

## model ev-a71 percentage as a function of time and age

model_viro_age_time <- gam(pos ~ s(age_adm) + s(time) + ti(age_adm,time),
                           family = binomial,method = "REML",data = df_viro)

collection_date_val <- seq(min(df_viro$time),
                           max(df_viro$time), le = 365)

new_data2 <- expand.grid(age_adm = age_val,
                         time = as.numeric(collection_date_val))

new_data2$pred <- predict(model_viro_age_time, newdata = new_data2, type = "response")
```


```{r}
cases_fit <- fit_gau_bs_ch1 |>  
    predict(newdata = df_cases_count_ch1_23, se.fit = TRUE) |> 
    as.tibble() |> 
    cbind(df_cases_count_ch1_23) |>
    mutate(
      # c_lwr = fit_gau_bs_ch1$family$linkinv(fit - 1.96 * se.fit),
      # c_upr = fit_gau_bs_ch1$family$linkinv(fit + 1.96 * se.fit),
      c_fit = fit_gau_bs_ch1$family$linkinv(fit)
           ) |> 
    dplyr::select(-c(se.fit,fit))

recontructed_df <- model_viro_age_time |>
  predict(newdata = df_cases_count_ch1_23 |> rename(age_adm = age), se.fit = TRUE) |>
  as.tibble() |>
  cbind(df_cases_count_ch1_23) |>
  mutate(v_lwr = model_viro_age_time$family$linkinv(fit - 1.96 * se.fit),
         v_upr = model_viro_age_time$family$linkinv(fit + 1.96 * se.fit),
         v_fit = model_viro_age_time$family$linkinv(fit)) |>
  dplyr::select(-c(se.fit,fit)) |>
  full_join(cases_fit, by = join_by(age, time,n)) |>
  mutate(ev71_c = c_fit*v_fit,
         ev71_u = c_fit*v_upr,
         ev71_l = c_fit*v_lwr)

recontructed_cases_count_df <- recontructed_df |>
  mutate(day = as.Date(time),
         week = as.Date(floor_date(day, "week"))) |> 
  group_by(week) |>
  summarise(cases = sum(c_fit),
            # cases_u = sum(c_upr),
            # cases_l = sum(c_lwr),
            ev71_cases = sum(ev71_c),
            ev71_cases_u = sum(ev71_u),
            ev71_cases_l = sum(ev71_l)) 
```


::: {.panel-tabset .nav-pills}

#### Fig 3A

```{r}
fig3a <- recontructed_df |>
  mutate(
    day = as.Date(time),
    week = as.Date(floor_date(day, "week")),
    period = case_when(
      week <= as.Date("2023-04-30") ~ "12/2022 - 4/2023",
      week > as.Date("2023-04-30") &
        week <= as.Date("2023-08-31") ~ "4/2023 - 8/2023",
      week > as.Date("2023-08-31") ~ "8/2023 - 12/2023"
    )
  ) |>
  group_by(period, age) |>
  summarise(
    ev_a71_cases = sum(ev71_c),
    raw_n = sum(n),
    .groups = "drop"
  ) |>
  ggplot(aes(x = age, y = ev_a71_cases)) +
  geom_col(alpha = 0.3) +
  geom_line(data = expected_incidence, aes(x = cohort, y = exp_inc),linewidth = 1) +
  geom_ribbon(
    data = expected_incidence,
    aes(x = cohort, y = exp_inc, ymin = lo_exp_inc, ymax = hi_exp_inc),
    alpha = 0.2,
    fill = "blue"
  ) +
  coord_cartesian(ylim = c(0, 2500)) +
  scale_x_continuous(position = "top") +
  facet_wrap( ~ period) +
  labs(y = "Number of cases", x = "Age (years)", tag = "A") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 18),
    plot.tag = element_text(face = "bold", size = 18),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    legend.title = element_text(size = 18),
    strip.text = element_text(size = 18),
    panel.grid.minor.x = element_blank()
  )
```


#### Fig 3B

```{r}
fig3b <- recontructed_cases_count_df |>
  ggplot(aes(x = week))+
  geom_line(aes(y = cases),linetype = "dashed")+
  # geom_ribbon(aes(ymin = cases_u,ymax = cases_l),fill = "blue",alpha = .3)+
  geom_line(aes(y = ev71_cases))+
  geom_ribbon(aes(ymin = ev71_cases_u,ymax = ev71_cases_l),fill = "red",alpha = .3)+
  geom_col(data = df_cases_count_ch1_23 |>
             mutate(day = as.Date(time),
                    week = as.Date(floor_date(day, "week")))|>
             group_by(week) |>
             summarise(cases = sum(n)),
           aes(x = week, y  = cases),alpha = .3)+
  geom_point(data = df_viro, aes(x = admission_date, y = pos*900),shape = "|",size = 2)+
  scale_x_date(date_breaks = "1 month",
               date_labels = "%b",
               name = "Collection date",
               limits = c(as.Date("2023-01-01"),as.Date("2023-12-31"))) +
  scale_y_continuous(limits = c(0,900),breaks = seq(0,900, le = 4))+
  labs(x = "Admission day (2023)",y = "Cases", tag = "B") +
  theme_minimal()+
  theme(axis.title.y = element_text(size = 18),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "bottom",
        plot.tag = element_text(face = "bold", size = 18),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 18),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 18))+
  guides(fill=guide_colourbar(barwidth=20),
         color = "none")
```


#### Fig 3C

```{r}
fig3c <- df_viro %>%
  group_by(adm_month) %>%
  count(sero_gr) %>%
  mutate(total = sum(n), per = n / total) %>%
  ggplot(aes(
    x = adm_month,
    y = per,
    fill = factor(sero_gr, levels = c("neg", "Other", "CV-A6", "EV", "EV-A71"))
  )) +
  geom_bar(position = "fill", stat = "identity") +
  geom_text(
    aes(label = total, y = 1),
    stat = "unique",
    vjust = -0.5,
    size = 8
  ) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b",
    limits = c(as.Date("2023-01-01"), as.Date("2023-12-31")),
    name = "Collection month (2023)"
  ) +
  scale_y_continuous(labels = scales::label_percent()) +  # extra space for label
  scale_fill_manual(
    values = c(
      "neg"    = "#00B0F6",
      "Other"  = "#E76BF3",
      "CV-A6"  = "#00BF7D",
      "EV"     = "#E59866",
      "EV-A71" = "#F8766D"
    )
  ) +
  labs(y = "Percentage (%)", fill = "Serotype group", tag = "C") +
  theme_minimal() +
  theme(
    axis.title.y  = element_text(size = 18),
    axis.title.x  = element_text(size = 18),
    axis.ticks.x  = element_blank(),
    legend.position = "bottom",
    plot.tag      = element_text(face = "bold", size = 18),
    axis.text.x   = element_text(size = 18),
    axis.text.y   = element_text(size = 18),
    legend.text   = element_text(size = 15),
    legend.title  = element_text(size = 18)
  )
```


:::

```{r}
#| fig-width: 20
#| fig-height: 16
#| out-width: "100%"

fig3c <- fig3c + coord_cartesian(clip = "off")

fig3 <- (fig3a /
           fig3b /
           fig3c) +
  plot_layout(heights = c(1, 1, 1))

ggsave(
  paste0(path2ms_fig,"fig3.svg"),
  plot = fig3,
  width = 20,
  height = 16,
  dpi = 500
)

fig3
```


## Supplementary fig

### Sup fig 1 

```{r}
#| fig-width: 15
#| fig-height: 9
#| out-width: "100%"

sup1
```

### Sup fig 2: geography

```{r}
#| fig-width: 12
#| fig-height: 12
#| out-width: "100%"

sup2 <- df23 %>%
  group_by(adm_week, district) %>%
  count() %>%
  ggplot(aes(x = adm_week, y = n)) +
  geom_col() +
  scale_x_date(
    date_breaks = "2 months",
    date_labels = "%b",
    name = "Admission week",
    limits = c(as.Date("2023-01-01"), as.Date("2023-12-31"))
  ) +
  theme_bw() +
  labs(y = "Number of admission") +
  facet_wrap( ~ district, ncol = 4) +
  theme(
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    axis.ticks.x = element_blank(),
    legend.position = "none",
    plot.tag = element_text(face = "bold", size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    strip.text.x = element_text(size = 18)
  )

ggsave(
  paste0(path2ms_fig,"sup2.svg"),
  plot = sup2,
  width = 12,
  height = 12,
  dpi = 500
)

sup2
```

### Sup fig 3

```{r}
sup3a <- df_cases_ch1_23 %>%
  group_by(adm_week, wave) %>%
  count() %>%
  ggplot(aes(x = adm_week, y = n, fill = wave)) +
  geom_col(alpha = 0.6, color = "black") +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %Y",
    name = "Collection date",
    limits = c(as.Date("2023-01-01"), as.Date("2023-12-31"))
  ) +
  scale_y_continuous(limits = c(0, 900), breaks = seq(0, 900, le = 4)) +
  scale_fill_manual(values = c("#582C83FF", "#FFC72CFF")) +
  theme_minimal() +
  labs(y = "Cases", tag = "A") +
  theme(
    axis.title.y = element_text(size = 18),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    plot.tag = element_text(face = "bold", size = 18),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = 18)
  ) +
  guides(fill = guide_colourbar(barwidth = 20), color = "none")
```

```{r}
#| fig-width: 20
#| fig-height: 16
#| out-width: "100%"

sup3 <- (sup3a /
           sup3b) +
  plot_layout(heights = c(1, 1))

ggsave(
  paste0(path2ms_fig,"sup3.svg"),
  plot = sup3,
  width = 20,
  height = 16,
  dpi = 500
)

sup3
```


### Sup fig 4: Correspondence between spatial distribution of serum samples and hopital catchment of CH1

::: {.panel-tabset .nav-pills}

#### Hospital catchment

```{r}
hos_catch <- df_cases_ch1_23 |>
  group_by(district) |>
  count() |>
  ungroup() |>
  mutate(per = n / sum(n)) |>
  right_join(qhtp, by = join_by(district == varname_2)) |>
  ggplot() +
  geom_sf(aes(fill = per, geometry = geom), show.legend = T) +
  paletteer::scale_fill_paletteer_c(
    "ggthemes::Classic Blue",
    na.value = "white",
    labels = scales::label_percent(),
    name = "Percentage"
  )+
  geom_sf(
    data = tdnd1,
    shape = 17,
    color = "yellow",
    size = 5
  ) +
  labs(tag = "A")+
  theme_void() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        plot.tag = element_text(face = "bold", size = 18))+
  guides(fill = guide_colourbar(barwidth = 15), color = "none")
```

#### Blood samples distribution

```{r}
sero_dis <- df_sero |>
  group_by(district) |>
  count() |>
  ungroup() |>
  mutate(per = n / sum(n)) |>
  right_join(qhtp, by = join_by(district == varname_2)) |>
  ggplot() +
  geom_sf(aes(fill = per, geometry = geom), show.legend = T) +
  paletteer::scale_fill_paletteer_c(
    "ggthemes::Classic Red",
    na.value = "white",
    labels = scales::label_percent(),
    name = "Percentage"
  ) +
  geom_sf(
    data = tdnd1,
    shape = 17,
    color = "yellow",
    size = 5
  ) +
  theme_void() +
  labs(tag = "B")+
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        plot.tag = element_text(face = "bold", size = 18))+
  guides(fill = guide_colourbar(barwidth = 15), color = "none")
```

#### Swabs sample spatial distribution

```{r}
viro_dis <- df_viro |>
  mutate(district = case_when(
    district %in% c("thu duc", "9", "2") ~ "thu duc",
    .default = district
  )) |> 
  group_by(district) |> 
  count() |> 
  ungroup() |>
  mutate(per = n / sum(n)) |>
  right_join(qhtp, by = join_by(district == varname_2)) |>
  ggplot() +
  geom_sf(aes(fill = per, geometry = geom), show.legend = T) +
  paletteer::scale_fill_paletteer_c(
    "ggthemes::Classic Green",
    na.value = "white",
    labels = scales::label_percent(),
    name = "Percentage"
  )+
  geom_sf(
    data = tdnd1,
    shape = 17,
    color = "yellow",
    size = 5
  ) +
  labs(tag = "C")+
  theme_void() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        plot.tag = element_text(face = "bold", size = 18))+
  guides(fill = guide_colourbar(barwidth = 15), color = "none")
```

:::

```{r}
#| fig-width: 12
#| fig-height: 6
#| out-width: "100%"

sup4 <- hos_catch | sero_dis | viro_dis

ggsave(
  paste0(path2ms_fig,"sup4.svg"),
  plot = sup4,
  width = 12,
  height = 6,
  dpi = 500
)

sup4

```

## Statistical analysis

### Fig 2A: case report

::: {.panel-tabset .nav-pills}

#### 12/22 - 4/23

```{r}
dat %>%
  bind_rows(.id = "id") %>%
  filter(age1 > 0) |>
  group_by(id, age) |>
  count() |> 
  ungroup() |> 
  filter(id == 1) |> 
  mutate(per = n*100/sum(n)) |> 
  as.data.frame()
```



#### 4/23 - 8/23

```{r}
dat %>%
  bind_rows(.id = "id") %>%
  filter(age1 > 0) |>
  group_by(id, age) |>
  count() |> 
  ungroup() |> 
  filter(id == 2) |> 
  mutate(per = n*100/sum(n))|> 
  as.data.frame()
```

#### 8/23 - 12/23

```{r}
dat %>%
  bind_rows(.id = "id") %>%
  filter(age1 > 0) |>
  group_by(id, age) |>
  count() |> 
  ungroup() |> 
  filter(id == 3) |> 
  mutate(per = n*100/sum(n))|> 
  as.data.frame()
```

:::

### Fig 2A: expected incidence

::: {.panel-tabset .nav-pills}

#### 12/22 - 4/23

```{r}
expected_incidence |> 
  filter(id == 1) |> 
  select(id,cohort,exp_inc,hi_exp_inc,lo_exp_inc) |> 
  summarise(median = median(exp_inc),
            q1 = quantile(exp_inc, 0.25),
            q3 = quantile(exp_inc, 0.75)) |> 
  as.data.frame()
```

Those age groups have the expected incidence higher than the median are considered as highly contribute

```{r}
expected_incidence |> 
  select(id,cohort,exp_inc,hi_exp_inc,lo_exp_inc) |> 
  filter(id == 1 & cohort <= 9) |> 
  summarise(mean = median(exp_inc),
            sd = sd(exp_inc)) |> 
  as.data.frame()
```


#### 04/23 - 08/23

```{r}
expected_incidence |> 
  select(id,cohort,exp_inc,hi_exp_inc,lo_exp_inc) |> 
  filter(id == 2 & cohort <= 1 ) |> 
  summarise(mean = mean(exp_inc),
            sd = sd(exp_inc)) |> 
  as.data.frame()
```


#### 08/23 - 12/23

find the mean expected ev-a71 infections

```{r}
expected_incidence |> 
  select(id,cohort,exp_inc,hi_exp_inc,lo_exp_inc) |> 
  filter(id == 3) |> 
  summarise(mean = mean(exp_inc)) |> 
  as.data.frame()
```

those age group has the expected incidence higher then mean is considered mainly contribute to the outbreak

```{r}
expected_incidence |> 
  select(id,cohort,exp_inc,hi_exp_inc,lo_exp_inc) |> 
  filter(id == 3 & exp_inc >= 94.89) |> 
  pull(cohort) |> 
  range()
```

caculate the mean and sd of those age group

```{r}
expected_incidence |> 
  select(id,cohort,exp_inc,hi_exp_inc,lo_exp_inc) |> 
  filter(id == 3 & cohort <= 7 & cohort >= 2) |> 
  summarise(mean = mean(exp_inc),
            sd = sd(exp_inc)) |> 
  as.data.frame()
```


:::

### Fig 2B

```{r}
fig2b_df <- outttt |>
  bind_rows(.id = "id") |>
  mutate(
    col_date = as.Date(collection_time),
    col_lab = format(col_date, format = "%b %Y"),
    col_lab = factor(col_lab, levels = c(
      "Dec 2022", "Apr 2023", "Aug 2023", "Dec 2023"
    ))
  )
```

::: {.panel-tabset .nav-pills}

#### Dec 22

Seroprevalence in children under 5 years old

```{r}
fig2b_df |>
  filter(id == 1 & cohort <= 5) |>
  reframe(across(c(fit, lwr, upr), \(x) range(x) * 100)) |>
  as.data.frame()
```

Seroprevalence in children higher than 5 years old

```{r}
fig2b_df |>
  filter(id == 1 & cohort > 5) |>
  reframe(across(c(fit, lwr, upr), \(x) range(x) * 100)) |> 
  as.data.frame()
```

#### April 2023

Find in which age group the sp increase

```{r}
expected_incidence |>
  filter(id == 1) |>
  select(cohort, sp_gap) |>
  mutate(
    prev_sign  = lag(sign(sp_gap)),
    sign_change = sign(sp_gap) != prev_sign,
    direction  = case_when(
      sign_change & sp_gap < 0 ~ "turned negative",
      sign_change & sp_gap > 0 ~ "turned positive"
    )
  ) |>
  filter(sign_change)
```

Which mean that the seroprevalence increased in children under 11.5 years old and over 13.5 years old

find range

```{r}
list(
  above = fig2b_df |> filter(id == 1 & cohort >= 13.5),
  below = fig2b_df |> filter(id == 1 & cohort <= 11.5)
) |>
  map(\(df) df |>
    reframe(across(c(fit, lwr, upr), \(x) range(x) * 100)) |>
    mutate(across(everything(), \(x) format(round(x, 2), nsmall = 2)))
  )
```

#### August 2023

Find in which age group the sp increase

```{r}
expected_incidence |>
  filter(id == 2) |>
  select(cohort, sp_gap) |>
  mutate(
    prev_sign  = lag(sign(sp_gap)),
    sign_change = sign(sp_gap) != prev_sign,
    direction  = case_when(
      sign_change & sp_gap < 0 ~ "turned negative",
      sign_change & sp_gap > 0 ~ "turned positive"
    )
  ) |>
  filter(sign_change)
```

Just select the under 8.5 years old age group

```{r}
fig2b_df |>
  filter(id == 3 & cohort <= 8.5) |>
  reframe(across(c(fit, lwr, upr), \(x) range(x) * 100)) |> 
  as.data.frame()
```

#### Dec 23

```{r}
expected_incidence |>
  filter(id == 3) |>
  select(cohort, sp_gap) |>
  mutate(
    prev_sign  = lag(sign(sp_gap)),
    sign_change = sign(sp_gap) != prev_sign,
    direction  = case_when(
      sign_change & sp_gap < 0 ~ "turned negative",
      sign_change & sp_gap > 0 ~ "turned positive"
    )
  ) |>
  filter(sign_change)
```

```{r}
fig2b_df |>
  filter(id == 4 & cohort <= 8.5) |>
  reframe(across(c(fit, lwr, upr), \(x) range(x) * 100)) |> 
  as.data.frame()
```



:::
