# Forecasting model

## Data path

```{r}
path2data <- paste0(
  "/Users/nguyenpnt/Library/CloudStorage/OneDrive-OxfordUniversityClinicalResearchUnit/Data/HFMD/cleaned/"
)

# Name of cleaned HFMD data file
cases_file <- "hfmd_hcdc.rds"

df_cases <- paste0(path2data, cases_file) |>
  readRDS()
```

## Library

Required packages:

```{r}
required <- c("BART",
              "tidyverse",
              "forecast",
              "MASS",
              "magrittr",
              "parallel",
              "lubridate",
              "tseries")
```

Installing those that are not installed yet:

```{r}
to_install <- required[!required %in% installed.packages()[, "Package"]]
if (length(to_install))
  install.packages(to_install)
```

Loading the packages for interactive use:

```{r}
invisible(sapply(required, library, character.only = TRUE))
```

## Function

plot bart function

```{r}
model_plot_w <- function(data) {
  data |>
    ggplot(aes(x = adm_week)) +
    geom_point(
      data = data |> dplyr::select(-week_run),
      aes(y = n),
      alpha = 0.4,
      size = .2
    ) +
    geom_line(aes(y = pred_t)) +
    geom_ribbon(
      aes(ymin = y_lower95, ymax = y_upper95),
      fill = "#4A90D9",
      alpha = 0.15,
      na.rm = TRUE
    ) +
    # Shaded 80% credible interval
    geom_ribbon(
      aes(ymin = y_lower80, ymax = y_upper80),
      fill = "#4A90D9",
      alpha = 0.25,
      na.rm = TRUE
    ) +
    scale_x_date(name = "Admission week") +
    scale_y_continuous(name = "Total cases") +
    labs(title = "Prediction for 3 weeks ahead")+
    theme_bw() +
    facet_wrap(~ week_run, ncol = 6)
}
```

function to run bart model to forecast 2023 data

```{r}

cores <- detectCores() - 1

# run_bart_month <- function(w, hfmd_data,y,lambda) {
# 
#   # hfmd_data <- hfmd_1324
#   # y = 2023
#   # w = 39
# 
#   train_data <- hfmd_data |>
#     filter(year < y | (year == y & week == w)) |>
#     data.frame() |>
#     na.omit()
# 
#   test_data <- hfmd_data |>
#     filter(year == y & week %in% c(w+1,w+2,w+3)) |>
#     data.frame()
# 
#   if (nrow(test_data) == 0)
#     return(NULL)
# 
#   model <- wbart(
#     x.train = train_data[, c("adm_week", "lag_t.1","lag_t.2","lag_t.3")],
#     y.train = train_data$y_transformed,
#     # x.test = test_data[, c("adm_week", "lag_t.1", "lag_t.2")],
#     ntree   = 1000,
#     ndpost  = 10000,
#     nskip   = 1000,
#     a       = 0.95,
#     b       = 2
#   )
# 
#   pred_draws <- predict(model,test_data[, c("adm_week", "lag_t.1", "lag_t.2", "lag_t.3")])
#   pred_ori <- InvBoxCox(pred_draws,lambda)
# 
#   result <- test_data |>
#     mutate(
#       pred_t = colMeans(pred_ori),
#       y_lower95  = apply(pred_ori, 2, quantile, 0.025),
#       y_upper95  = apply(pred_ori, 2, quantile, 0.975),
#       y_lower80  = apply(pred_ori, 2, quantile, 0.10),
#       y_upper80  = apply(pred_ori, 2, quantile, 0.90),
#       week_run = w,
#       year_run = y
#     )
# 
#   return(result)
# }
```

Function to run ARIMA model

```{r}
run_arima_week <- function(w, y, hfmd_data, lambda) {
  
  # i        = 1
  # w         = year_week_grid$w[i]
  # y         = year_week_grid$y[i]
  # hfmd_data = hfmd_1324
  # lambda    = lambda
  
  
  # Expanding window training data
  train_data <- hfmd_data |>
    dplyr::filter(year < y | (year == y & week < w)) |>
    data.frame() |>
    na.omit()
  
  # Test = 3 weeks ahead from week w
  test_data <- hfmd_data |>
    dplyr::filter(year == y & week >= w & week <= w + 2) |>
    data.frame()
  
  if (nrow(test_data) == 0 | nrow(train_data) < 10) return(NULL)
  
  # Convert training to time series object
  ts_train <- ts(train_data$y_transformed, frequency = 52)
  
  # Fit auto ARIMA with tryCatch
  model <- auto.arima(ts_train)
  
  
  if (is.null(model)) return(NULL)
  
  # Fix 1: forecast exactly nrow(test_data) steps, not always 3
  h  <- nrow(test_data)
  fc <- forecast(model, h = h, level = c(80, 95))
  
  # Fix 2: safely back-transform with InvBoxCox instead of manual formula
  # handles edge cases like lambda = 0 (log transform)
  result <- test_data |>
    mutate(
      pred_t    = as.numeric(InvBoxCox(fc$mean,          lambda)),
      y_lower95 = as.numeric(InvBoxCox(fc$lower[, "95%"], lambda)),
      y_upper95 = as.numeric(InvBoxCox(fc$upper[, "95%"], lambda)),
      y_lower80 = as.numeric(InvBoxCox(fc$lower[, "80%"], lambda)),
      y_upper80 = as.numeric(InvBoxCox(fc$upper[, "80%"], lambda)),
      horizon   = row_number(),
      week_run  = w,
      year_run  = y
    )
  
  return(result)
}
```

## Analysis

```{r}
hfmd_1324 <- df_cases %>%
  mutate(adm_week = as.Date(floor_date(admission_date, "week"))) %>%
  group_by(adm_week) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(
    x = 1:nrow(.),
    week = week(adm_week),
    month = month(adm_week),
    year = year(adm_week)
  ) %>%
  rename(time = x)

```

I applied Box-cox transformation to normalize the count data

```{r}
bc <- boxcox(n ~ adm_week,data = hfmd_1324)
lambda <- bc$x[which.max(bc$y)]
```

Add transformed cases and lag 1-week term

```{r}
hfmd_1324 %<>%
  mutate(
    y_transformed = ((n^lambda - 1) / lambda),
    lag_t.1 = lag(y_transformed,1),
    lag_t.2 = lag(y_transformed,2),
    lag_t.3 = lag(y_transformed,3)
  )
```

Before transform:

```{r}
hfmd_1324 |>
  ggplot(aes(n))+
  geom_histogram(binwidth = 100)
```

After transform:

```{r}
hfmd_1324 |>
  ggplot(aes(y_transformed))+
  geom_histogram(binwidth = .5)
```

## BART model

First, I used data from 2013 to 2022 for training and using 2023 for testing. Then I combined 2013-2022 data with each lag of 3 weeks in 2023 to see the prediction 3 weeks ahead

```{r}
# year_week_grid <- expand.grid(y = c(2023), w = 1:52)
# 
# results_list <- mclapply(
#   X        = 1:nrow(year_week_grid),
#   FUN      = function(i) {
#     run_bart_month(
#       w         = year_week_grid$w[i],
#       y         = year_week_grid$y[i],
#       hfmd_data = hfmd_1324
#     )
#   },
#   mc.cores = cores
# )

# pred_3w <- bind_rows(results_list)
# save(pred_3w,file = "pred_3w.RData")

load("./data/pred_3w.RData")
```

week 1-26, 2023

```{r}
#| fig-width: 17
#| fig-height: 12
#| out-width: "100%"
pred_3w |>
  filter(week_run <= 26) |> 
  model_plot_w()
```

week 27-52, 2023

```{r}
#| fig-width: 17
#| fig-height: 12
#| out-width: "100%"
pred_3w |>
  filter(week_run > 26) |> 
  model_plot_w()
```

week 1-24, 2024

```{r}
# year_week_grid <- expand.grid(y = 2024, w = 1:26)
# 
# results_list <- mclapply(
#   X        = 1:nrow(year_week_grid),
#   FUN      = function(i) {
#     run_bart_month(
#       w         = year_week_grid$w[i],
#       y         = year_week_grid$y[i],
#       hfmd_data = hfmd_1324,
#       lambda = lambda
#     )
#   },
#   mc.cores = cores
# )
# pred_3w_24 <- bind_rows(results_list)
# save(pred_3w_24, file = "pred_3w_24.RData")
```


```{r}
#| fig-width: 17
#| fig-height: 12
#| out-width: "100%"

load("./data/pred_3w_24.RData")

pred_3w_24 |>
  filter(week_run <= 24) |> 
  model_plot_w()
```


## ARIMA model

ARIMA assume data is stationary. A time series is considered stationary if its statistical properties, such as mean and variance, remain constant over time.

I used adf and kpss test for stationary testing of hfmd time-series data. For ARIMA model, I used data from 2012-2022 for training.

```{r}
train_data <- hfmd_1324 |>
    dplyr::filter(year < 2023 | (year == 2023 & week < 1)) |>
    data.frame() |>
    na.omit()
```

First I used number of admission for stationary testing. There are 2 test for stationary testing

-   adf test, null hypothesis: The time series has a unit root, meaning it is non-stationary.

-   kpss test, null hypothesis: The series is stationary

Create a time-series

```{r}
ts_train <- ts(train_data$n, frequency = 52)
```

adf test

```{r}
adf.test(ts_train) ## can reject the null hypothesis
```

kpss test

```{r}
kpss.test(ts_train) ## can reject the null hypothesis
```

Usually, there should be a contrast between those 2 test for stationary. If a data is stationary when p-value of adf test \< 0.05 and kpss test \> 0.05. So the results above state the non-stationary, then I applied the box-cox transformation for the time-series data and retest.

```{r}
ts_train_t <- ts(train_data$y_transformed, frequency = 52)
adf.test(ts_train_t) 
kpss.test(ts_train_t) 
```

So I used the box-cox transformation for model

```{r}
# # Explicit cluster setup
# cl <- makeCluster(cores)
# 
# # Export everything the workers need
# clusterExport(cl, varlist = c("run_arima_week", "hfmd_1324",
#                               "lambda", "year_week_grid"))
# 
# # Load packages on each worker
# clusterEvalQ(cl, {
#   library(forecast)
#   library(dplyr)
# })
# 
# # Run
# sarima_results_list <- parLapply(
#   cl  = cl,
#   X   = 1:nrow(year_week_grid),
#   fun = function(i) {
#     tryCatch(
#       run_arima_week(
#         w         = year_week_grid$w[i],
#         y         = year_week_grid$y[i],
#         hfmd_data = hfmd_1324,
#         lambda    = lambda
#       ),
#       error = function(e) return(NULL)
#     )
#   }
# )
# 
# # Always stop cluster when done
# stopCluster(cl)
# 
# results_arima <- bind_rows(sarima_results_list)
# save(results_arima,file = "results_arima.RData")

load("./data/results_arima.RData")

```

Week 1-26, 2023

```{r}
#| fig-width: 17
#| fig-height: 12
#| out-width: "100%"

results_arima |>
  filter(week_run <= 26) |>
  model_plot_w()
```

Week 27-52, 2023

```{r}
#| fig-width: 17
#| fig-height: 12
#| out-width: "100%"

results_arima |>
  filter(week_run > 26) |>
  model_plot_w()
```

Week 1-24, 2024

```{r}
#| fig-width: 17
#| fig-height: 12
#| out-width: "100%"

# results_arima_24 <- bind_rows(sarima_results_list)
# 
# save(results_arima_24,file = "results_arima_24.RData")

load("./data/results_arima_24.RData")

results_arima_24 |> 
  filter(week_run <= 24) |> 
  model_plot_w()
```

