---
bibliography: references.bib
---

```{r}
library(gtsummary)
library(dplyr)
library(stringi)
library(ggplot2)
library(lubridate)
library(tidyr)
library(paletteer)
invisible(Sys.setlocale("LC_TIME", "English"))
```


# Virology data description

```{r}
library(readxl)
viro_df <- read_excel("D:/OUCRU/hfmd/data/03EI Data 2023 shared.xlsx")
```

## Demographic

```{r}
viro2 <- viro_df %>% 
  mutate(city = City %>% 
           trimws(which = "both") %>% 
           stri_trans_general("latin-ascii") %>% 
           tolower(),
         sero_gr = case_when(
           SeroGroup1 == "ENT" ~ "EV",
           SeroGroup1 != "ENT" ~ SeroGroup1
         ),
         admission_date = as.Date(DateAdmission),
         adm_month = month(admission_date),
         age_adm = interval(DateBirth, admission_date) / years(1)) %>% 
  select(city,sero_gr,admission_date,adm_month,age_adm,DateBirth)

viro2$city <- factor(viro2$city,
                         levels = c("tp hcm",viro2$city[!viro2$city == "tp hcm"] %>% 
                                      unique()))

viro2 %>% 
  tbl_summary(by=sero_gr,
              include = c(city,sero_gr))
```

```{r,results='hide'}
link <- "https://data.opendevelopmentmekong.net/dataset/999c96d8-fae0-4b82-9a2b-e481f6f50e12/resource/2818c2c5-e9c3-440b-a9b8-3029d7298065/download/diaphantinhenglish.geojson?fbclid=IwAR1coUVLkuEoJRsgaH81q6ocz1nVeGBirqpKRBN8WWxXQIJREUL1buFi1eE"

vn_spatial <- sf::st_read(link) %>% mutate(
  city = 
    case_when(
      Name == "TP. Ho Chi Minh" ~ "tp hcm",
      Name != "TP. Ho Chi Minh" ~ Name
    ) %>% 
    trimws(which = "both") %>% 
    stri_trans_general("latin-ascii") %>% 
    tolower()
)

viro2 %>% 
  filter(sero_gr == "EV-A71") %>% 
  group_by(city,sero_gr) %>% 
  count() %>% 
  ungroup() %>% 
  full_join(.,vn_spatial, by = join_by(city)) %>% 
  ggplot() + 
  geom_sf(aes(fill = n,geometry = geometry))+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    na.value="white",
                                    name = "Numbers of EV-A71 samples")+
  theme_void()+
  theme(legend.position="bottom")
```

## Time 

```{r}
viro2 %>% 
  mutate(adm_month = as.Date(floor_date(admission_date, "month"))) %>%
  group_by(adm_month,sero_gr) %>% 
  count() %>% 
  ggplot(aes(x = adm_month,y=n,fill = sero_gr))+
  geom_col()+
  scale_x_date(date_labels = "%b",
               date_breaks = "1 month")+
  labs(fill = "Sero group", y = "Number of samples",x = "Month of admission")+
  theme_minimal()+
  theme(legend.position="bottom")
```

## Age

```{r}
viro2 %>%
  mutate(adm_month = as.Date(floor_date(admission_date, "month")),
         age_bin = cut(age_adm,
                       breaks = seq(0, max(age_adm, na.rm = TRUE) + 0.5, by = 0.5),
                       right = FALSE)) %>%
  na.omit() %>%
  group_by(age_bin, sero_gr) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(perc = n / sum(n)) %>%
  ungroup() %>%
  complete(age_bin,sero_gr,fill = list(perc=0)) %>% 
  ggplot(aes(x = age_bin, y = perc, fill = sero_gr)) +
  geom_col(position = "stack", color = "black") +
  ggsci::scale_fill_jco() +
  labs(x = "Age (years)", y = "Percentage (%)", fill = "Serotype") +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Age and time

### EV-A71 positive samples as funtion of age and time

```{r}
#| fig-width: 10
#| fig-height: 6 
#| out-width: "100%"

viro2 %>%
  mutate(adm_month = month(admission_date,label = TRUE),
         adm_month2 = as.Date(floor_date(admission_date, "month"))%>% as.character(),
         age_bin = cut(age_adm,
                       breaks = seq(0, max(age_adm, na.rm = TRUE) + 0.5, by = 0.5),
                       right = FALSE)) %>%
  na.omit() %>% 
  filter(sero_gr == "EV-A71") %>%
  select(adm_month2,age_adm) %>%
  as.data.frame() %>% 
  ggplot(aes(x=adm_month2, y = age_adm)) +
  stat_density(
    aes(fill = after_stat(count)),
    geom = "raster",
    position = "identity",
    interpolate = TRUE
  )+
  scale_fill_paletteer_c("grDevices::Inferno")+
  theme_minimal()+
  scale_y_reverse(name = "Age (years)",breaks = seq(0,12,by = 1))+
  labs(fill = "Number of EV-A71 samples",x = "Month of admission")+
  theme(legend.position = "bottom",
        legend.key.width =  unit(1, "cm"))

```

### age distribution EV-A71 positive samples as funtion of age and time

```{r}
#| fig-width: 10
#| fig-height: 6 
#| out-width: "100%"

viro2 %>%
  mutate(adm_month = month(admission_date,label = TRUE),
         adm_month2 = as.Date(floor_date(admission_date, "month"))%>% as.character(),
         age_bin = cut(age_adm,
                       breaks = seq(0, max(age_adm, na.rm = TRUE) + 0.5, by = 0.5),
                       right = FALSE)) %>%
  na.omit() %>% 
  filter(sero_gr == "EV-A71" & month(adm_month2) >= 6) %>%
  select(adm_month2,age_adm) %>%
  as.data.frame() %>% 
  ggplot(aes(x=adm_month2, y = age_adm)) +
  stat_density(
    aes(fill = after_stat(density)),
    geom = "raster",
    position = "identity",
    interpolate = TRUE
  )+
  scale_fill_paletteer_c("grDevices::Inferno")+
  theme_minimal()+
  scale_y_reverse(name = "Age (years)",
                  limits = c(0,6),
                  breaks = seq(0,6,by = 1))+
  labs(fill = "Density",x = "Month of admission")+
  theme(legend.position = "bottom",
        legend.key.width =  unit(1, "cm"))

```

I limited the age from 0-6 years old

### Select from May-Aug

This age profile coresspond with 2 sucessive collection time points in the 2nd figure in Fig1A 

```{r}
#| fig-width: 10
#| fig-height: 6 
#| out-width: "100%"

viro2 %>%
  mutate(adm_month = month(admission_date,label = TRUE),
         adm_month2 = as.Date(floor_date(admission_date, "month"))%>% as.character(),
         age_bin = cut(age_adm,
                       breaks = seq(0, max(age_adm, na.rm = TRUE) + 0.5, by = 0.5),
                       right = FALSE)) %>%
  na.omit() %>% 
  filter(sero_gr == "EV-A71" & month(adm_month2) <= 8) %>%
  select(adm_month2,age_adm) %>%
  as.data.frame() %>% 
  ggplot(aes(x=adm_month2, y = age_adm)) +
  stat_density(
    aes(fill = after_stat(count)),
    geom = "raster",
    position = "identity",
    interpolate = TRUE
  )+
  scale_fill_paletteer_c("grDevices::Inferno")+
  theme_minimal()+
  scale_y_reverse(name = "Age (years)",
                  limits = c(0,6),
                  breaks = seq(0,6,by = 1))+
  labs(fill = "Number of EV-A71 samples",x = "Month of admission")+
  theme(legend.position = "bottom",
        legend.key.width =  unit(1, "cm"))
```

Raw data

```{r}
count <- viro2 %>%
  mutate(adm_month = month(admission_date,label = TRUE),
         adm_month2 = as.Date(floor_date(admission_date, "month"))%>% as.character(),
         age_bin = cut(age_adm,
                       breaks = seq(0, max(age_adm, na.rm = TRUE) + 0.5, by = 0.5),
                       right = FALSE)) %>%
  na.omit() %>% 
  filter(sero_gr == "EV-A71" & month(adm_month2) <= 8) %>% 
  ggplot(aes(age_adm))+
  geom_histogram(binwidth = 0.5,
                 color = "white",fill = "black",alpha = 0.5)+
  scale_x_continuous(breaks = seq(0,12,by = 1),
                     limits = c(0,12))+
  labs(x = "Age(years)",y = "Number of EV-A71 samples")+
  theme_minimal()


per <- viro2 %>%
  mutate(adm_month = month(admission_date,label = TRUE),
         adm_month2 = as.Date(floor_date(admission_date, "month"))%>% as.character(),
         age_bin = cut(age_adm,
                       breaks = seq(0, max(age_adm, na.rm = TRUE) + 0.5, by = 0.5),
                       right = FALSE)) %>%
  na.omit() %>%
  filter(month(adm_month2) <= 8) %>%
  group_by(age_bin,sero_gr) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(perc = n / sum(n)) %>%
  ungroup() %>% 
  complete(age_bin,sero_gr,fill = list(perc =0)) %>%
  filter(sero_gr == "EV-A71") %>% 
  ggplot(aes(x = age_bin, y = perc))+
  geom_col()+
  scale_y_continuous(name = "Percentage of EV-A71 samples",
                     labels = scales::label_percent(scale = 100))+
  labs(x = "Age group")+
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)
  )

library(patchwork)

count/
  per

```

### Select from Aug-Dec

This age profile coresspond with 2 sucessive collection time points in the 3rd figure in Fig1A 

```{r}
#| fig-width: 10
#| fig-height: 6 
#| out-width: "100%"

viro2 %>%
  mutate(adm_month = month(admission_date,label = TRUE),
         adm_month2 = as.Date(floor_date(admission_date, "month"))%>% as.character(),
         age_bin = cut(age_adm,
                       breaks = seq(0, max(age_adm, na.rm = TRUE) + 0.5, by = 0.5),
                       right = FALSE)) %>%
  na.omit() %>% 
  filter(sero_gr == "EV-A71" & month(adm_month2) >= 8) %>%
  select(adm_month2,age_adm) %>%
  as.data.frame() %>% 
  ggplot(aes(x=adm_month2, y = age_adm)) +
  stat_density(
    aes(fill = after_stat(count)),
    geom = "raster",
    position = "identity",
    interpolate = TRUE
  )+
  scale_fill_paletteer_c("grDevices::Inferno")+
  theme_minimal()+
  scale_y_reverse(name = "Age (years)",
                  limits = c(0,6),
                  breaks = seq(0,6,by = 1))+
  labs(fill = "Number of EV-A71 samples",x = "Month of admission")+
  theme(legend.position = "bottom",
        legend.key.width =  unit(1, "cm"))
```

Raw data

```{r}
count <- viro2 %>%
  mutate(adm_month = month(admission_date,label = TRUE),
         adm_month2 = as.Date(floor_date(admission_date, "month"))%>% as.character(),
         age_bin = cut(age_adm,
                       breaks = seq(0, max(age_adm, na.rm = TRUE) + 0.5, by = 0.5),
                       right = FALSE)) %>%
  na.omit() %>% 
  filter(sero_gr == "EV-A71" & month(adm_month2) >= 8) %>% 
  ggplot(aes(age_adm))+
  geom_histogram(binwidth = 0.5,
                 color = "white",fill = "black",alpha = 0.5)+
  scale_x_continuous(breaks = seq(0,12,by = 1),
                     limits = c(0,12))+
  labs(x = "Age(years)",y = "Number of EV-A71 samples")+
  theme_minimal()


per <- viro2 %>%
  mutate(adm_month = month(admission_date,label = TRUE),
         adm_month2 = as.Date(floor_date(admission_date, "month"))%>% as.character(),
         age_bin = cut(age_adm,
                       breaks = seq(0, max(age_adm, na.rm = TRUE) + 0.5, by = 0.5),
                       right = FALSE)) %>%
  na.omit() %>%
  filter(month(adm_month2) >= 8) %>%
  group_by(age_bin,sero_gr) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(perc = n / sum(n)) %>%
  ungroup() %>% 
  complete(age_bin,sero_gr,fill = list(perc =0)) %>% 
  filter(sero_gr == "EV-A71") %>% 
  ggplot(aes(x = age_bin, y = perc))+
  geom_col()+
  scale_y_continuous(name = "Percentage of EV-A71 samples",
                     labels = scales::label_percent(scale = 100))+
  labs(x = "Age group")+
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)
  )

library(patchwork)

count/
  per

```


## Epidemiology analysis

![Figure1](D:/OUCRU/assigned github/hfmd/fig_manuscript/fig2_3.svg)
